### 2023-07-17 fig/convergence.tex
init

\begin{tikzpicture}
    \begin{semilogyaxis}[
        xlabel={\# samples},
        ylabel={relative error},
        legend entries={$\varphi_3$, $\varphi_2$, $i_\mathrm{L}$},
        % legend entries={$\varphi_3$, $\varphi_2$},
        legend pos=north east
        ]

        % index 1
        \addplot[color=TUDa-1a, mark=*, mark size=3pt] table {do1_convergence_phi_3.dat};
        \addplot[color=TUDa-9a, mark=triangle*, mark size=3pt] table {do1_convergence_phi_2.dat};
        \addplot[color=TUDa-6a, mark=square*, mark size=3pt] table {do1_convergence_i_L.dat};

        % index 2
        % \addplot[color=TUDa-1a, mark=*, mark size=3pt] table {do2_convergence_phi_3.dat};
        % \addplot[color=TUDa-9a, mark=triangle*, mark size=3pt] table {do2_convergence_phi_2.dat};
    \end{semilogyaxis}
\end{tikzpicture}