### 2023-11-30 doc/fig/ial.tex

% \begin{tikzpicture}
%     \node[rectangle, rounded corners, draw=black, semithick] (dae) at (0,0) {DAE};
%     \node[rectangle, rounded corners, draw=black, semithick] (ode) at (-2,-2) {\textcolor{TUDa-1a}{ODE}};
%     \node[rectangle, rounded corners, draw=black, semithick] (ae) at (2,-2) {\textcolor{TUDa-9a}{AE}};
%     \node[rectangle, rounded corners, draw=black, semithick] (dd) at (-2,-4) {\textcolor{TUDa-1a}{Differential DOFs}};
%     \node[rectangle, rounded corners, draw=black, semithick] (ad) at (2,-7) {\textcolor{TUDa-9a}{Algebraic DOFs}};

%     \node[circle, fill=TUDa-9a, minimum size=3pt, inner sep=0] (ddae) at (2,-4) {};

%     \draw[->, >=latex, semithick, color=black] (dae) -- (ode) node[above, midway, sloped] {decouple};
%     \draw[->, >=latex, semithick, color=black] (dae) -- (ae) node[above, midway, sloped] {decouple};
%     \draw[<-, >=latex, semithick, color=TUDa-1a, dashed] (dd) -- (ode) node[above, midway, sloped] {learn};

%     \draw[-, semithick, color=TUDa-9a, dotted] (dd) -- (ddae);
%     \draw[-, semithick, color=TUDa-9a, dotted] (ae) -- (ddae);
%     \draw[<-, >=latex, semithick, color=TUDa-9a, dotted] (ad) -- (ddae) node[above, midway, sloped] {reconstruct};
% \end{tikzpicture}

% landscape
\begin{tikzpicture}
    \node[rectangle, rounded corners, draw=black, semithick] (dae) at (0,0) {DAE};
    \node[rectangle, rounded corners, draw=black, semithick] (ode) at (2,-2) {\textcolor{TUDa-1a}{ODE}};
    \node[rectangle, rounded corners, draw=black, semithick] (ae) at (2,2) {\textcolor{TUDa-9a}{AE}};
    \node[rectangle, rounded corners, draw=black, semithick] (dd) at (5.5,-2) {\textcolor{TUDa-1a}{Differential DOFs}};
    \node[rectangle, rounded corners, draw=black, semithick] (ad) at (9.5,2) {\textcolor{TUDa-9a}{Algebraic DOFs}};

    \node[circle, fill=TUDa-9a, minimum size=3pt, inner sep=0] (ddae) at (5.5,2) {};

    \draw[->, >=latex, semithick, color=black] (dae) -- (ode) node[above, midway, sloped] {decouple};
    \draw[->, >=latex, semithick, color=black] (dae) -- (ae) node[above, midway, sloped] {decouple};
    \draw[<-, >=latex, semithick, color=TUDa-1a, dashed] (dd) -- (ode) node[above, midway, sloped] {learn};

    \draw[-, semithick, color=TUDa-9a, dotted] (dd) -- (ddae);
    \draw[-, semithick, color=TUDa-9a, dotted] (ae) -- (ddae);
    \draw[<-, >=latex, semithick, color=TUDa-9a, dotted] (ad) -- (ddae) node[above, midway, sloped] {reconstruct};
\end{tikzpicture}